---
output: pdf_document
---
Sampling Distributions
========================================================
### Math 141
### Week 7

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```


```{r}
library(tidyverse)
library(pdxTrees)
library(infer)
```

## Important Notes

* To construct a Sampling Distribution for a statistic, we need access to the entire population so that we can take repeated samples from the population. 

* But if we have access to the entire population, then we don't need a sampling distribution!

* Often we only have access to a single sample.  In these cases, we will need to approximate the sampling distribution. 

* We are studying the properties of a sampling distribution so that we can figure out how to approximate it well!


## Data

Let's use the `pdxTrees` data and focus on all the trees in a specific park: Mt Tabor Park.

```{r}
trees_mt_tabor <- get_pdxTrees_parks(park = "Mt Tabor Park")
```

### Examples 


#### Example 1: Sampling Distribution for the Proportion of Douglas-Fir


```{r}
# Add column for Douglas-Fir
trees_mt_tabor <- trees_mt_tabor %>%
  mutate(Douglas_Fir = case_when(
    Common_Name == "Douglas-Fir"  ~ "yes",
    Common_Name != "Douglas-Fir"  ~ "no"))

# Population distribution
ggplot(data = trees_mt_tabor, mapping = aes(x = Douglas_Fir)) +
  geom_bar(aes(y = ..prop.., group = 1), stat = "count") 

#True population parameter
summarize(trees_mt_tabor, parameter = mean(Douglas_Fir == "yes"))
```



```{r}
# Draw random samples
samples <- trees_mt_tabor %>% 
  rep_sample_n(size = 50, reps = 4)
  
# Graph the samples
ggplot(data = samples, mapping = aes(x = Douglas_Fir)) +
  geom_bar(aes(y = ..prop.., group = 1), stat = "count") +
  facet_wrap( ~ replicate)
```

```{r}
# Construct the sampling distribution
samp_dist <- trees_mt_tabor %>% 
  rep_sample_n(size = 50, reps = 1000) %>%
  group_by(replicate) %>%
  summarize(statistic = mean(Douglas_Fir == "yes"))

# Graph the sampling distribution
ggplot(data = samp_dist, mapping = aes(x = statistic)) +
  geom_histogram(bins = 21)
```

**Observations**:

* Shape?
* Center?
* Spread?

```{r}
mean_of_means <- summarize(samp_dist, mean(statistic)) %>%
  pull()
mean_of_means

std_error <- summarize(samp_dist, sd(statistic)) %>%
  pull()
std_error

```

**Questions of Interest:**

1. What happens to the center/spread/shape as we increase the sample size?

+ Median approaches mean
+ Spread decreses


To investigate, re-run the code but change the sample size (`size`) to 100, 150.

2. What happens to the center/spread/shape if the true parameter changes?

To investigate re-run the code but change "Douglas-Fir" to "BigLeaf Maple".

#### Example 2: Sampling Distribution for the Mean Tree Height

```{r}
# Population distribution
ggplot(data = trees_mt_tabor, mapping = aes(x = Tree_Height)) +
  geom_histogram()

#True population parameter
summarize(trees_mt_tabor, parameter = mean(Tree_Height))
```



```{r}
# Draw random samples
samples <- trees_mt_tabor %>% 
  rep_sample_n(size = 50, reps = 4)
  
# Graph the samples
ggplot(data = samples, mapping = aes(x = Tree_Height)) +
  geom_histogram() +
  facet_wrap( ~ replicate)
```

```{r}
# Construct the sampling distribution
samp_dist <- trees_mt_tabor %>% 
  rep_sample_n(size = 50, reps = 1000) %>%
  group_by(replicate) %>%
  summarize(statistic = mean(Tree_Height))

# Graph the sampling distribution
ggplot(data = samp_dist, mapping = aes(x = statistic)) +
  geom_histogram(bins = 21)
```

**Observations**:

* Shape?
* Center?
* Spread?

```{r}
mean_of_means <- summarize(samp_dist, mean(statistic)) %>%
  pull()
mean_of_means

std_error <- summarize(samp_dist, sd(statistic)) %>%
  pull()
std_error

```


**Questions of Interest:**

1. What happens to the center/spread/shape as we increase the sample size?

To investigate, re-run the code but change the sample size (`size`) to 100, 150.

2. How does the shape of the population distribution compare to the shape of the sampling distribution? 
